---
title: "Data 607 HW 3"
author: "Kayleah Griffen"
date: "2023-02-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(readr)
library(tidyverse)
library(RCurl)
```

## Question 1

Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS".

```{r data-stats}
x <- getURL("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv")
y <- read.csv(text = x)
str_subset(y$Major, "DATA") 
str_subset(y$Major, "STATISTICS")
```

## Question 2

Write code that transforms the data below:
[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"

Into a format like this:
c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r}
strStart = '[1] "bell pepper" "bilberry"   "blackberry"  "blood orange"
[5] "blueberry"  "cantaloupe"  "chili pepper" "cloudberry"
[9] "elderberry"  "lime"     "lychee"    "mulberry"
[13] "olive"    "salal berry"'

temp <- str_replace_all(strStart, "[^[:alpha:]]+", " ")
temp <- str_trim(temp, side="both")
temp <- str_split_fixed(temp, "\\s", n=Inf)
temp <- c(temp)
print(temp) # As a vector
# From here I looked up how to make the vector print comma separated
# https://stackoverflow.com/questions/6347356/creating-a-comma-separated-vector
cat("c(", paste(shQuote(temp, type="cmd"), collapse=", "), ")") # Formatted
```

## Question 3

Describe, in words, what these expressions will match:

1. (.)\1\1

The `()` indicates a group, the `.` is a meta-character where the `.` will match any character. So `(.)` is group 1 and will match any character. The `\1` is a backreference, referencing group 1. So `\1\1` indicates the same text matched by the first group should be matched again, twice, because there are two `\1`s. The full `(.)\1\1` indicates the same character should appear three times in a row to be matched.

```{r}
ts <- c("abc", "aabc", "aaabbc", "123", "12223")
str_subset(ts, "(.)\\1\\1")
```

2. "(.)(.)\\2\\1"

The `(.)(.)` indicates two groups, both containing all characters. The `\\2` is a backreference to group 2 and `\\1` is a backreference to group 1. The full `(.)(.)\\2\\1` indicates any character (say `x`) followed by another character (say `y`), followed by `y` followed by `x` should match. If `x=y` then any 4 of the same characters in a row will match. 

```{r}
ts <- c("aaaaa", "abba", "aaba", "abbbbb", "xyyx", "111222333")
str_subset(ts, "(.)(.)\\2\\1")
```

3. (..)\1

The `(..)` indicates group 1 is any two characters, the `\1` indicates those same two characters should be repeated for a match. So for example `abab` would match.

```{r}
ts <- c("aaaaa", "abab", "aaba", "abbbbb", "12213", "111222333")
str_subset(ts, "(..)\\1")
```

4. "(.).\\1.\\1"

The `(.)` indicates group 1 is any character, the `.` is any character, the `\\1` refers to group 1. So in this case for a match the character in group 1, must be followed by a character, followed by group 1s character, followed by any character, followed by group 1s character again.

```{r}
ts <- c("aaaaa", "abaca", "aaba", "12131", "12213", "111222333")
str_subset(ts, "(.).\\1.\\1")
```

5. "(.)(.)(.).*\\3\\2\\1"

The three `(.)` indicates three groups of any character. The `.` indicates any character. The `*` is another meta character letting a pattern be optional or repeat any number of times. So the `.*` indicates a character repeated any number of times, including zero times is an option. The `\\3\\2\\1` means that group 3, then group 2, then group 1 should be repeated to complete the match. So to get a match from `(.)(.)(.).*\\3\\2\\1` the following must be true: character 1, then character 2, then character 3, then any character or no character, then character 3, then character 2, then character 1. 

```{r}
ts <- c("abczzzzzcba", "aaaaaa", "abcdefg", "123321", "12213", "111222333")
str_subset(ts, "(.)(.)(.).*\\3\\2\\1")
```

## Question 4

Construct regular expressions to match words that:

1. Start and end with the same character.

```{r}
ts <- c("Hannah", "hello", "kayak", "aa", "121", "111222333")
str_subset(ts, regex("^(.).*\\1$", ignore_case = T))
```

2. Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

```{r}
ts <- c("Church", "abab", "abc", "123321", "1212", "111222333")
str_subset(ts, regex("([A-Za-z][A-Za-z]).*\\1", ignore_case = T))
```

3. Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

```{r}
ts <- c("Eleven", "aaaaaa", "abcdefg", "12121", "12213", "111222333")
str_subset(ts, regex("([A-Za-z]).*\\1.*\\1", ignore_case = T))
```
